---
title: "BT1101 Practice Exam"
author: "BT1101 Student"
date: '29 April 2022, 9:00 - 11:30AM'
output: html_document
---

## Instructions

-   **Rename your R Markdown file `FE_Student#.rmd`**, replacing Student# with your Student number that starts with A (e.g. A0202020) and the output will automatically be `FE_Student#.html`.

-   Select `output: html_document`.

-   Include all code chunks, so include `echo=TRUE` in all chunks.

-   Replace the placeholder text, "Type your answer here.", with your own.

-   Pre-install and include any `library('package_name')` statements before exam starts. Remember that there is NO Internet connection during the exam.

-   Code and type your answer in this Rmarkdown file during exam but **keep in mind that you need to copy and paste all your answer (r-chunk and text) into Essay Answer section for each question in Examplify.**

-   Submit your both R Markdown file (.rmd) and HTML (.html) to the folder "Final Exam Submission" in Luminus after Examplify submission and Internet reconnection.

-   This Rmarkdown file serves as a reference. **Only answers submitted in Examplify will be graded.** Zero point will be given for blank submission in Examplify even if you have submitted a complete Rmarkdown/HTML file.

## Preparation

```{r preparation, echo=TRUE, warning = FALSE, message = FALSE}

# predictive/prescriptive analytics
library(tseries) 
library(forecast)
library(lpSolve)
library(TTR)

# descriptive analytics
library(psych)
library(Rmisc)
library(rcompanion)
library(rpivotTable)
library(EnvStats) 
library(car)
library(rstatix)

# general use
library(wooldridge)
library(dplyr)
library(tidyr)
library(knitr)
```

## Question 2: Course Selection

Here is an example of a constraints table that you can put into R Markdown:

| Maximize total profit using decision variables $X_1$, $X_2$ | Profit = 0.15 $X_1$ + 0.40 $X_2$ |
|-------------------------------------------------------------|----------------------------------|
| Subject to                                                  |                                  |
| Budget Constraint                                           | 0.20$X_1$ + 0.70$X_2$ $\leq$ 100 |
| Space Constraint                                            | $X_1$ + $X_2$ $\leq$ 200         |
| Non-Negativity Constraint 1                                 | $X_1$ + $\quad$ $\geq$ 0         |
| Non-Negativity Constraint 2                                 | $\quad$ + $X_2$ $\geq$ 0         |

Natalie has decided to switch her career to data science, as she feels it has more prospects than her previous industry. She is eyeing a "pico-masters" program at her local university, where she has to complete 40 units of courses to satisfy the pico-degree.

The program offers the following courses, in STatistics, ProGramming, and Data Management, along with their associated costs and pre-requisites. The pre-requisites for each course must be fulfilled before students are allowed to take that course.

Natalie has also indicated her personal interest in each course.

| Course | Units | Pre-requisites | Cost    | Interest |
|:-------|:------|:---------------|:--------|:---------|
| ST1    | 10    | \-             | \$1,000 | 8        |
| ST2    | 10    | ST1            | \$1,500 | 4        |
| ST3    | 10    | ST2            | \$2,000 | 6        |
| PG1    | 10    | \-             | \$1,000 | 7        |
| PG2a   | 10    | PG1            | \$1,500 | 5        |
| PG2b   | 10    | PG1            | \$1,500 | 6        |
| PG3    | 10    | PG2a or PG2b   | \$2,000 | 3        |
| DM1    | 10    | \-             | \$1,000 | 4        |
| DM2    | 10    | DM1            | \$1,500 | 6        |
| DM3    | 10    | DM2            | \$2,000 | 7        |

#### Q2.(a) How would you write out the following constraint as a linear constraint? (1 mark)

**"Students must take ST1 before ST2"**

We define $X_1$ and $X_2$ to represent Natalie taking ST1 and ST2 courses respectively.

$X_1$ $\geq$ $X_2$

#### Q2.(b) Suppose that her only goal is to minimize cost and satisfy the requirements of the degree.

-   Identify the decision variables, objective function, and the relevant constraints.
-   Write them out in a table. You may keep the "Pre-requisite" constraints in the form X + Y \<= Z or X + Y \>= Z or X + Y = Z for this table.

Do not solve the problem yet. (3 marks)

Define Decision Variables: $X_1$, $X_2$, $X_3$, $X_4$, $X_5$, $X_6$, $X_7$, $X_8$, $X_9$ and $X_{10}$ corresponding to whether or not she takes the ten courses in the respective order as seen in the table above

| Minimize Cost using decision variables = $X_1$, $X_2$, $X_3$, $X_4$, $X_5$, $X_6$, $X_7$, $X_8$, $X_9$, $X_{10}$ | Cost = 1000 $X_1$ + 1500 $X_2$ + 2000 $X_3$ + 1000 $X_4$ + 1500 $X_5$ + 1500 $X_6$ + 2000 $X_7$ + 1000 $X_8$ + 1500 $X_9$ + 2000 $X_{10}$ |
|------------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------------------------------------------------------------------|
| **Set of Constraints**                                                                                           | Subject to                                                                                                                                |
| 40 Units of Courses Requirements                                                                                 | 10 $X_1$ + 10 $X_2$ + 10 $X_3$ + 10 $X_4$ + 10 $X_5$ + 10 $X_6$ + 10 $X_7$ + 10 $X_8$ + 10 $X_9$ + 10 $X_{10}$ $\ge$ 40                   |
| Pre-Requisite for ST2                                                                                            | $X_1$ $\geq$ $X_2$                                                                                                                        |
| Pre-Requisite for ST3                                                                                            | $X_2$ $\geq$ $X_3$                                                                                                                        |
| Pre-Requisite for PG2a                                                                                           | $X_4$ $\geq$ $X_5$                                                                                                                        |
| Pre-Requisite for PG2b                                                                                           | $X_4$ $\geq$ $X_6$                                                                                                                        |
| Pre-Requisite for PG3                                                                                            | $X_5$ + $X_6$ $\geq$ $X_7$                                                                                                                |
| Pre-Requisite for DM2                                                                                            | $X_8$ $\geq$ $X_9$                                                                                                                        |
| Pre-Requisite for DM3                                                                                            | $X_9$ $\geq$ $X_{10}$                                                                                                                     |
| Binary, Integer, Non-Negativity Constraints                                                                      | $X_1$ to $X_{10}$ all binary, integers and $\geq 0$.                                                                                      |

#### Q2.(c) Natalie just read a condition she missed out earlier. In order to finish the PicoMasters, she needs to finish a specialization in one of the three tracks, which is fulfilled by completing the "Level 3" version of that course. How will you specify this as a linear constraint? (2 marks)

$X_3$ + $X_7$ + $X_{10}$ $\geq$ 1

#### Q2.(d) Natalie admitted that "cost minimization" should be a lesser priority to "interest maximization". Rewrite the problem to maximize her interest. Note that since cost is still a factor for her, she decides to take EXACTLY 40 units, no more than that.

-   Write out the table again, including the new constraint from (c) and the changes in the preceding paragraph.
-   If you are not confident of going straight from this table to R, please take the additional step of converting the linear constraints into a "fully spelled-out" constraint with all the 0s, 1s and −1s, as we did in lecture.
-   Using R, write a linear program to solve Natalie's problem.

Which courses should she pick? (7 marks)

| Maximize Interest using decision variables = $X_1$, $X_2$, $X_3$, $X_4$, $X_5$, $X_6$, $X_7$, $X_8$, $X_9$, $X_{10}$ | Interest = 8 $X_1$ + 4 $X_2$ + 6 $X_3$ + 7 $X_4$ + 5 $X_5$ + 6 $X_6$ + 3 $X_7$ + 4 $X_8$ + 6 $X_9$ + 7 $X_{10}$     |
|----------------------------------------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------|
| **Set of Constraints**                                                                                               | Subject to                                                                                                          |
| 40 Units of Courses Requirements                                                                                     | 10 $X_1$ + 10 $X_2$ + 10 $X_3$ + 10 $X_4$ + 10 $X_5$ + 10 $X_6$ + 10 $X_7$ + 10 $X_8$ + 10 $X_9$ + 10 $X_{10}$ = 40 |
| Pre-Requisite for ST2                                                                                                | $X_1$ $\geq$ $X_2$                                                                                                  |
| Pre-Requisite for ST3                                                                                                | $X_2$ $\geq$ $X_3$                                                                                                  |
| Pre-Requisite for PG2a                                                                                               | $X_4$ $\geq$ $X_5$                                                                                                  |
| Pre-Requisite for PG2b                                                                                               | $X_4$ $\geq$ $X_6$                                                                                                  |
| Pre-Requisite for PG3                                                                                                | $X_5$ + $X_6$ $\geq$ $X_7$                                                                                          |
| Pre-Requisite for DM2                                                                                                | $X_8$ $\geq$ $X_9$                                                                                                  |
| Pre-Requisite for DM3                                                                                                | $X_9$ $\geq$ $X_{10}$                                                                                               |
| Specialization Required                                                                                              | $X_3$ + $X_7$ + $X_{10}$ $\geq$ 1                                                                                   |
| Binary, Integer, Non-Negativity Constraints                                                                          | $X_1$ to $X_{10}$ all binary, integers and $\geq 0$.                                                                |

```{r q2d, echo = TRUE}

#defining parameters
objective.fn <- c(8, 4, 6, 7, 5, 6, 3, 4, 6, 7)
const.mat <- matrix(c(10, 10, 10, 10, 10, 10, 10, 10, 10, 10,
                      1, -1, 0, 0, 0, 0, 0, 0, 0, 0, # Pre-Requisite for ST2
                      0, 1, -1, 0, 0, 0, 0, 0, 0, 0, # Pre-Requisite for ST3
                      0, 0, 0, 1, -1, 0, 0, 0, 0, 0, # Pre-Requisite for PG2a
                      0, 0, 0, 1, 0, -1, 0, 0, 0, 0, # Pre-Requisite for PG2b
                      0, 0, 0, 0, 1, 1, -1, 0, 0, 0, # Pre-Requisite for PG3
                      0, 0, 0, 0, 0, 0, 0, 1, -1, 0, # Pre-Requisite for DM2
                      0, 0, 0, 0, 0, 0, 0, 0, 1, -1, # Pre-Requisite for DM3
                      0, 0, 1, 0, 0, 0, 1, 0, 0, 1 # Specialization
                      ), 
                      ncol=10 , byrow=TRUE) 
const.dir <- c("=", rep(">=", 8))
const.rhs <- c(40, rep(0, 7), 1)

#solving model
lp.solution <- lp("max", objective.fn, const.mat, const.dir, const.rhs, binary.vec = c(1:10))
lp.solution$solution #decision variables values 
lp.solution
```

As seen above, the optimal solution is $X_1$ = 1, $X_2$ = 1, $X_3$ = 1, $X_4$ = 1, $X_5$ = 0, $X_6 = 0$, $X_7$ = 0, $X_8$ = 0, $X_9$ = 0 and $X_{10}$ = 0. This means that she should pick ST1, ST2, ST3 and PG1 courses in order to maximize interest. As seen above, the objective function is 25 which means that the optimal solution would give her 25 units of interest.
